---
title: "challenge_fas"
format: html
editor: visual
---

## Coding Challenge

------------------------------------------------------------------------

Simple example grouped bar plot

```{r, echo = FALSE}
##example grouped bar plot
 # library
library(ggplot2)
 
# create a dataset
specie <- c(rep("sorgho" , 3) , rep("poacee" , 3) , rep("banana" , 3) , rep("triticum" , 3) )
condition <- rep(c("normal" , "stress" , "Nitrogen") , 4)
value <- abs(rnorm(12 , 0 , 15))
data <- data.frame(specie,condition,value)
 
# Grouped
ggplot(data, aes(fill=condition, y=value, x=specie)) + 
    geom_bar(position="dodge", stat="identity")
```

------------------------------------------------------------------------

### Review of data:

```{r}

library(readr)

#bps_aoi_attributes
bps_aoi_attributes <- read_csv("input_data/bps_aoi_attributes.csv")
head(bps_aoi_attributes)
tail(bps_aoi_attributes)

#bps_model_number_name
bps_model_number_name <- read_csv("input_data/bps_model_number_name.csv")
head(bps_model_number_name)
tail(bps_model_number_name)
## seems useless

#combine_raw
combine_raw <- read_csv("input_data/combine_raw.csv")
head(combine_raw)
tail(combine_raw)

#LF16_BPS_200
LF16_BPS_200 <- read_csv("input_data/LF16_BPS_200.csv")
head(LF16_BPS_200)
tail(LF16_BPS_200)
## also may be useless?

#rec_con_modified
ref_con_modified <- read_csv("input_data/ref_con_modified.csv")
head(ref_con_modified)
tail(ref_con_modified)

#scls_aoi_attributes
scls_aoi_attributes <- read_csv("input_data/scls_aoi_attributes.csv")
head(scls_aoi_attributes)
tail(scls_aoi_attributes)

```

------------------------------------------------------------------------

### Plan:

bps_aoi_attributes - need Freq, BPS_NAME, BPS_MODEL, and VALUE

combine_raw - present succession classes per BpS

ref_con_modified - past reference percentages for each succession class per BpS -- pivot longer?

scls_aoi_attributes - names to merge

------------------------------------------------------------------------

### Need to:

1.  clean up data - just whats needed above

-   i think i need to tidy ref_con_modified to fit like combine_raw

2.  tie in names with models for:

-   Ozark-Ouachita Dry-Mesic Oak Forest
-   Ozark-Ouachita Shortleaf Pine-Bluestem Woodland
-   Ozark-Ouachita Shortleaf Pine-Oak Forest and Woodland

3.  merge datasets with name, model, percentages for past and present, and value?

------------------------------------------------------------------------

### Fake plot:

```{r, echo = FALSE, results = FALSE, message = FALSE}
#load libraries
library(tidyverse)

#set seed for reproducibility
set.seed(123)

#create dataset
data <- data.frame(
  bps = rep(c("eco A", "eco B", "eco C"), each = 9),
  label = rep(c("A", "B", "C", "D", "E", "UN", "UE", "Agriculture", "Developed"), times = 3),
  ref_cur = rep(c("Present", "Past"), each = 27),
  amount = sample(0:60, 27, replace = TRUE)
)

data

#set x-axis order
level_order <- c("Developed", "Agriculture", "UE", "UN", "E", "D", "C", "B", "A")

# Plot
plot <-
  ggplot(data, aes(fill=(ref_cur), y=amount, x=label)) +
  geom_col(width = 0.8, position = position_dodge()) +
  coord_flip() +
  facet_grid(. ~bps) +
  labs(
    title = "Succession classes past and present",
    subtitle = "Fake chart with fake data",
    caption = "Data from landfire.gov",
    x = "",
    y = "Amount (units unknown)",
    fill = "Time period") +
  scale_x_discrete(limits = level_order) + 
  theme_light()

plot

```

------------------------------------------------------------------------


### Data wrangling:

```{r}
###
#select only variables needed from bps_aoi_attributes
bps_a_a <- subset(bps_aoi_attributes, select=c(BPS_NAME, BPS_MODEL, VALUE, Freq))

#only keep names needed
library(tidyverse)
bps <- bps_a_a %>% filter(BPS_NAME %in% c("Ozark-Ouachita Dry-Mesic Oak Forest", "Ozark-Ouachita Shortleaf Pine-Bluestem Woodland", "Ozark-Ouachita Shortleaf Pine-Oak Forest and Woodland" ))
###

###
#now in combine_raw -- filter only values 2160, 2132, and 2141 in Var1
# combine_raw %>% filter(Var1 %>% c("2160", "2132", "2141")) 
#Error in `filter()`:
# ℹ In argument: `Var1 %>% c("2160", "2132", "2141")`.
# Caused by error:
# ! `..1` must be of size 146 or 1, not size 149.

#trying to filter a different way
#first, subset columns needed
combine_raw_sub <- subset(combine_raw, select=c(Var1, Var2, Freq))
#make a vector of values needed
comb_raw <- 
  combine_raw_sub |> 
  filter(Var1 == 2160 | Var1 == 2132 | Var1 == 2141)
###

###
#now cleaning LF16_BPS_200 like bps_aoi_attributes
lf16 <- subset(LF16_BPS_200, select=c(BPS_NAME, BPS_MODEL, VALUE))

#keep only names needed
lf <- lf16 %>% filter(BPS_NAME %in% c("Ozark-Ouachita Dry-Mesic Oak Forest", "Ozark-Ouachita Shortleaf Pine-Bluestem Woodland", "Ozark-Ouachita Shortleaf Pine-Oak Forest and Woodland" ))
###

###
#filter ref_con_modified for specific model codes "15070_44", "13040_32_43_44_49", "13670_32_44"
rcm <- ref_con_modified %>% filter(Model_Code %in% c("15070_44", "13040_32_43_44_49", "13670_32_44"))

#pivot longer rcm data frame
rcm_long <- rcm %>% pivot_longer(
  cols = !Model_Code, 
  names_to = "label", 
  values_to = "score"
  )

#change column name to BPS_MODEL to match other datasets
colnames(rcm_long) <- c('BPS_MODEL', 'class', 'score')
###

###
#get succession classes from scls_aoi_attributes A, B, C, D, E, UN, UE, Agriculture, Developed
scls <- scls_aoi_attributes %>% filter(LABEL %in% c("A", "B", "C", "D","E","UN","UE", "Agriculture", "Developed"))
#may not use this, just manually add?
###

#

###






```



